### 2020-01-09 nlp_models.tex
some ebasic text added

\charles{more here ?}
We analyze GPT models deployed with the popular HuggingFace PyTorch library.
GPT has 12 layers, with 4 Multi-head Attention Blocksm, giving $48$ Layer Weight Matrices $\mathbf{W}$.
Each Block has 2 components, the Self Attention (attn) and the Projection (proj) matrics.  
The self-attention  matrices are larger, of dimension ($2304\times 768$) or ($3072\times 768$).
The projection layer concatenates the self-attention results into a vector (of dimension $768$).
This gives many large, typically sparse weight matrices, and they can be
 \emph{poorly correlated} when trained with insufficient data--as we shall see below.

@@

\charles{Since we show this table, maybe we should discuss GPT2 scaling models}

\begin{table}[t]
\small
\begin{center}
\begin{tabular}{|p{1in}|c|c|c|c|c|}
\hline
   &    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
 Series & \#Layers   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{\infty}$ & $\hat{\alpha}=\alpha\log\lambda_{max}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
\hline
 GPT & & & & \\
 GPT2 (small) & & & & \\
 GPT2-medium & & & & \\
 GPT2-large & & & & \\
 GPT2-xls & & & & \\

\hline
\end{tabular}
\end{center}
\caption{Average metrics for pretrainnd OpenAI GPT and GPT2 models.}
\label{table:models}
\end{table}

\paragraph{Heavy Tailed Power Law Exponents in GPT and GPT2}

@@

\paragraph{Correlation Flow in GPT and GPT2}

@@

>>>>

\charles{more here ?}
We analyze GPT models deployed with the popular HuggingFace PyTorch library.
GPT has 12 layers, with 4 Multi-head Attention Blocks, giving $48$ Layer Weight Matrices $\mathbf{W}$.
Each Block has 2 components, the Self Attention (attn) and the Projection (proj) matrics.  
The self-attention  matrices are larger, of dimension ($2304\times 768$) or ($3072\times 768$).
The projection layer concatenates the self-attention results into a vector (of dimension $768$).
This gives $50$ large, typically sparse weight matrices, and they can be
 \emph{poorly correlated} when trained with insufficient data--as we shall see below.

@@

The additional OpenAI GPT2 (English) models are: \emph{gpt-medium, got-large, and gpt-xl}, 
having include $12, 24, 36, and 48$ layers, resp., with increasingly larger weight matrices.
The model card for GPT2 is published on github.\footnote{\url{https://github.com/openai/gpt-2/blob/master/model_card.md}}.
Table \ref{table:nlp} reports results for the average log norm metrics, using \emph{weightwatcher (0.2.3)},
and with fully reproducible Jupyter notebooks.\footnote{\url{https://github.com/CalculatedContent/kdd2020}}

\begin{table}[t]
\small
\begin{center}
\begin{tabular}{|p{1in}|c|c|c|c|c|}
\hline
   &    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
 Series & \#Layers   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{\infty}$ & $\hat{\alpha}=\alpha\log\lambda_{max}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
\hline
 GPT & & & & \\
 GPT2 (small) & & & & \\
 GPT2-medium & & & & \\
 GPT2-large & & & & \\
 GPT2-xl & & & & \\

\hline
\end{tabular}
\end{center}
\caption{Average Log Norm Metrics for pretrainnd OpenAI GPT and GPT2 models.}
\label{table:nlp}
\end{table}

\paragraph{The Heavy Tailed Power Law Exponents in GPT and GPT2}

are very differenmt, with GPT2 having both a notably smaller mean $\alpha$, and far fewer, unusually large outliers.
Figure \ref{fig:gpt-alphs-hist} shows the empirical density (histogram) of $\alpha$
for all layers in GPT (blue) and GPT2 (red).  \charles{discuss more}

@@

\paragraph{The Correlation Flow in GPT and GPT2} also differs significantly between GPT and GPT2.
Figure \ref{fig:gpt-alpha-layer} plots $\alpha$ vs the layer id for each model.

\charles{Discuss Spectral Norm, alpha-Norm}

@@

Figure \ref{fig:gpt-alpha-layers} ...

For this series of GPT2 models, the average $\alpha$ still decreases with increasing model size,
althogh, the differences are less noticible than between the GPT and GPT2 models.
Unlike GPT, however, the Layer Spectral Norms $\vert\mathbf{W}\Vert_{\infty}$ 
and Alpha-Norms $\vert\mathbf{W}\Vert_{\alpha}^{\alpha}$
 behave  as expected for GPT2 layers, with the larger models 
consistently having  smaller norms.
